A survey of registration practices among observational researchers using preexisting datasets

### 2023-03-17 manuscript/manuscript.Rmd
small fixes to the plots and tables

keywords          : "keywords"
wordcount         : "X"
# bibliography      : "references.bib"
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

@@

```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=6}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
# processed %>% 
#   select(contains("typically_")) %>% 
#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>% 
#   nrow()

@@

# Create figure
typically_plot <- 
  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.7) +
  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )

@@

# Create figure
ecaw_plot <- 
  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 4.7) +
  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )

@@

```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4.5}
# Used methods figures
method_plot_data <-
  processed %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind the data analyst",
        type == "confirmatory" ~ "Contain confirmatory analysis",
        type == "exploratory" ~ "Contain exploratory analysis",
        type == "preregistered" ~ "Are preregistered",
        type == "script" ~ "Share analysis scripts",
      ),
      type = factor(type, levels = c(
        "Contain exploratory analysis",
        "Contain confirmatory analysis",
        "Blind the data analyst",
        "Share analysis scripts",
        "Are preregistered"
      ))
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

@@

method_plot_data %>% 
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  labs(
    x = "Percentage",
    fill = "Response",
    title = "The studies using preexisting observational data that I am involved in..."
  ) +
  papaja::theme_apa() +
  theme(
    axis.title.y = element_blank(),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  ) 
```

@@

```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4, message=FALSE}
# Check if there is any missing and don't understand responses separately
alspac_missing_text <-
  processed %>% 
  select(contains("alspac_")) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
    ) %>% 
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>% 
  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
  replace_na(list(value = "Missing")) %>% 
  group_by(item , value) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(text = paste(value, "=", n))

# Prepare plot data
alspac_plot_data <- 
  processed %>%
  select(contains("alspac_")) %>% 
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree")
          )
      )
    ) %>% 
  rename(
    `If ALSPAC required that I\nuse an ECAW, I would be less willing to\nuse their data in my research` = alspac_less_willing,
    `If ALSPAC ran a study\non ECAWs, I would opt-in.` = alspac_opt_in,
    `ALSPAC should run\na study on ECAWs.` = alspac_study,
    `I would prefer using\nan ECAW than using typical preregistration` = alspac_prefer_ecaw
  )

@@

# Create figure
alspac_plot <- 
  plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, group.order = names(alspac_plot_data), text.size = 6) +
  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 12),
    legend.title = element_blank()
  )

alspac_plot
```

@@

>>>>

keywords          : "keywords"
wordcount         : "X"
# bibliography      : "references.bib"
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
appendix: "supplementary_materials.Rmd"

classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

@@

```{r typicallyEcawPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=12.2, fig.height=6}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
# processed %>%
#   select(contains("typically_")) %>%
#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
#   nrow()

@@

# Create figure
typically_plot <- 
  plot(typically_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  labs(title = "Typically, studies that analyze preexisting observational datasets\n(such as the ALSPAC dataset) are...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank()
  )

@@

# Create figure
ecaw_plot <- 
  plot(ecaw_plot_data, digits = 1, ordered = FALSE, legend.position = "right", text.size = 5) +
  labs(title = "Compared to a typical study using preexisting observational data,\na study using an ECAW would be...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 13.5),
    legend.title = element_blank(),
    # legend.box.spacing = unit(0, "cm")
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -2, unit = "cm")
  )

@@

```{r methodPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5}
# Used methods figures
method_plot_data <-
  processed %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind the\ndata analyst",
        type == "confirmatory" ~ "Contain\nconfirmatory analysis",
        type == "exploratory" ~ "Contain\nexploratory analysis",
        type == "preregistered" ~ "Are preregistered",
        type == "script" ~ "Share analysis scripts",
      ),
      type = factor(type, levels = c(
        "Contain\nexploratory analysis",
        "Contain\nconfirmatory analysis",
        "Blind the\ndata analyst",
        "Share analysis scripts",
        "Are preregistered"
      ))
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

@@

method_plot_data %>% 
  # filter(scale %ni% c("Missing", "I don't understand the question")) %>% 
  ggplot() +
  aes(
    x = percentage,
    y = type,
    fill = scale
  ) +
  geom_bar(position = "stack", stat = "identity") +
  scale_x_continuous(
    limits = c(0, 1),
    labels = scales::label_percent()
  ) +
  scale_fill_viridis(option = "A", discrete = TRUE) +
  labs(
    x = "Percentage",
    fill = "Response",
    title = "The studies using preexisting observational data that I am involved in..."
  ) +
  papaja::theme_apa() +
  theme(
    title = element_text(size = 17),
    axis.title.y = element_blank(),
    axis.text = element_text(size = 15),
    axis.title.x = element_text(size = 17),
    legend.text = element_text(size = 14),
    legend.title = element_blank(),
    legend.box.margin = margin(t = 0, r = 0, b = 0, l = -0.5, unit = "cm")
  ) 
```

@@

```{r alspacPlot, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=14, fig.height=4, message=FALSE}
# Check if there is any missing and don't understand responses separately
alspac_missing_text <-
  processed %>% 
  select(contains("alspac_")) %>% 
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
    ) %>% 
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>% 
  mutate(item = str_to_title(str_remove(item , "^[^_]*_"))) %>% 
  replace_na(list(value = "Missing")) %>% 
  group_by(item , value) %>% 
  summarise(
    n = n()
  ) %>% 
  ungroup() %>% 
  mutate(text = paste(value, "=", n))

# Prepare plot data
alspac_plot_data <- 
  processed %>%
  select(contains("alspac_")) %>% 
  # Transform not wanted response values to NA
  # likert::likert drops NA values silently
  # When var transformed to factor these values would be transformed to NA
  # Automatically but I try to be explicit
  mutate(
    across(
      everything(),
      ~ case_when(
        . %in% c("I don't understand the question", "Unsure") ~ NA_character_,
        TRUE ~ .
      )
    )
  ) %>% 
  mutate(
    across(
      .fns = ~ factor(., levels = c(
          "Strongly disagree",
          "Somewhat disagree",
          "Neither agree nor disagree",
          "Somewhat agree",
          "Strongly agree")
          )
      )
    ) %>% 
  rename(
    `If ALSPAC required that I use an ECAW, I would be less willing to use their data in my research` = alspac_less_willing,
    `If ALSPAC ran a study on ECAWs, I would opt-in.` = alspac_opt_in,
    `ALSPAC should run a study on ECAWs.` = alspac_study,
    `I would prefer using an ECAW than using typical preregistration` = alspac_prefer_ecaw
  )

@@

# Create figure
plot(alspac_plot_likert_data, digits = 1, ordered = TRUE, group.order = names(alspac_plot_data), text.size = 6) +
  labs(title = "Thinking about a study you may run with ALSPAC data (or one that you have recently run)...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_text(size = 13),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.title = element_blank()
  )
```

@@

# (APPENDIX) Appendix {-}

```{r child = "supplementary_materials.Rmd"}
```

### 2023-03-17 manuscript/manuscript.tex
small fixes to the plots and tables

>>>>

\hypertarget{appendix-appendix}{%
\appendix}

\hypertarget{supplementary-materials-c}{%
\section{Supplementary materials C}\label{supplementary-materials-c}}

{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/numStudies-1}

\caption{Caption goes here...}\label{fig:numStudies}
\end{figure}

\begin{table}[H]

\caption{\label{tab:unnamed-chunk-7}What programming language or software do you use for your analyses of preexisting observational data?}
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrr}
\toprule
\textbf{Programming language} & \textbf{N} & \textbf{Percentage of respondents}\\
\midrule
R & 65 & 61\\
\cmidrule{1-3}
Stata & 48 & 45\\
\cmidrule{1-3}
SPSS & 17 & 16\\
\cmidrule{1-3}
SAS & 15 & 14\\
\cmidrule{1-3}
Python & 6 & 6\\
\cmidrule{1-3}
Missing & 4 & 4\\
\cmidrule{1-3}
Mplus & 3 & 3\\
\cmidrule{1-3}
Bash & 2 & 2\\
\cmidrule{1-3}
MATLAB & 1 & 1\\
\cmidrule{1-3}
Nextflow & 1 & 1\\
\cmidrule{1-3}
plink2 & 1 & 1\\
\bottomrule
\end{tabular}}
\end{table}

{\centering \includegraphics[width=1\linewidth]{manuscript_files/figure-latex/concerned-1}

\caption{Caption goes here...}\label{fig:concerned}
\end{figure}

### 2023-03-17 manuscript/supplementary_materials.Rmd
small fixes to the plots and tables

---
title             : "Supplementary materials: A survey of registration practices among observational researchers using preexisting datasets"
shorttitle        : "Supplementary materials"

keywords          : "keywords"
wordcount         : "X"
floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no
figurelist        : no
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library("papaja")
library(tidyverse)

# Source R scripts
r_scripts <- list.files(here::here("R/"), full.names = TRUE)
purrr::walk(r_scripts, source)

# Loading data
processed <- read_csv(here("data/processed/ecaw_processed_data.csv"))
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Table

```{r}
# programming languages table
processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  replace_na(list(lang = "Missing")) %>% 
  arrange(desc(n)) %>% 
  rename(
    `Programming language` = lang,
    `Number of occurance` = n
  ) %>% 
  kbl(
  .,
  format = "latex",
  col.names = c("Programming language", "Number of occurance"),
  booktabs = T,
  escape = F,
  caption = "Programming Languages Used to Analyse Preexisting Observational Data",
  centering = T,
  position = "H") %>%
  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
  row_spec(row = 0, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
```

# Figure

```{r}
# Create text for missing responses
n_studies_missing_text <- 
  processed %>% 
  filter(is.na(n_studies)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create histogram of the number of observational studies done
processed %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram() +
  annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
  labs(
    x = "Number of observational studies done prior using ALSPAC",
    y = "Count"
  ) +
  papaja::theme_apa()
```

>>>>

# Supplementary materials C

```{r numStudies, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=6, fig.height=3.2}
# Create text for missing responses
n_studies_missing_text <- 
  processed %>% 
  filter(is.na(n_studies)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

# Create histogram of the number of observational studies done
processed %>% 
  # Exclude missing responses
  filter(!is.na(n_studies)) %>% 
  ggplot() +
  aes(x = n_studies) +
  geom_histogram(binwidth = 3) +
  # annotate(geom = "text", x = 200, label = n_studies_missing_text, y = 30) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  labs(
    x = "Number of publications",
    y = "Number of respondents",
    title = "Approximately how many studies have you published using\na preexisting observational dataset (e.g., the ALSPAC dataset)?"
  ) +
  papaja::theme_apa()

# ggplot() +
#   geom_histogram(
#     data = dplyr::filter(valami, n_studies == 0),
#     mapping = aes(x = n_studies), binwidth = 1, color = 'red') +
#   geom_histogram(
#     data = dplyr::filter(valami, n_studies != 0),
#     mapping = aes(x = n_studies), binwidth = 5) +
#   # annotate(geom = "text", label = n_studies_missing_text, y = 30) +
#   scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
#   labs(
#     x = "Number of observational studies done prior using ALSPAC",
#     y = "Count"
#   ) +
#   papaja::theme_apa()
```

```{r}
# Programming languages table
# We use the number of all participants who responded to the questionnaire as the denominator
denominator <- nrow(processed)

processed %>% 
  rename(lang = programming_language) %>% 
  separate_rows(lang, sep = ",") %>% 
  count(lang) %>% 
  replace_na(list(lang = "Missing")) %>% 
  # Removing irrelevant
  filter(lang != "irrelevant") %>% 
  ungroup() %>% 
  mutate(
    `Percentage of respondents` = round(n / denominator * 100, 0)
  ) %>% 
  arrange(desc(n)) %>% 
  rename(
    `Programming language` = lang,
    `N` = n
  ) %>% 
  kbl(
  .,
  format = "latex",
  col.names = c("Programming language", "N", "Percentage of respondents"),
  booktabs = T,
  escape = F,
  caption = "(ref:table1c-caption)",
  centering = T,
  position = "H") %>%
  kable_styling(full_width = F,  position = "left", latex_options = "scale_down") %>%
  row_spec(row = 0, bold = TRUE) %>%
  collapse_rows(columns = 1, valign = "top", latex_hline = "major", row_group_label_position = "identity")
```

(ref:table1c-caption) What programming language or software do you use for your analyses of preexisting observational data?

```{r concerned, warning=FALSE, fig.cap="Caption goes here...", fig.align="center", out.width="100%", fig.width=10, fig.height=4}
# Concerned figure data preparation
concerned_plot_data <-
  processed %>% 
  select(concerned) %>% 
  mutate(
    concerned = tolower(concerned),
    concerned = factor(concerned, levels = c(
      "very much less concerned",
      "less concerned",
      "somewhat less concerned",
      "as concerned as a typical researcher in my field",
      "somewhat more concerned",
      "more concerned",
      "very much more concerned"
      )
    )
  )

# Check distinct values in data
# distinct(concerned_plot_data, concerned)

# Create missing values text
concerned_missing_text <- 
  concerned_plot_data %>% 
  filter(is.na(concerned)) %>% 
  summarise(
    n = n(),
    missing_text = paste0("missing = ", n)
  ) %>% 
  pull(missing_text)

concerned_plot_data <- filter(concerned_plot_data, !is.na(concerned))

# Create likert package data not including the missing values
concerned_likert_plot_data <- likert(concerned_plot_data)

# Create figure
plot(concerned_likert_plot_data, digits = 1, text.size = 6, group.order = names(concerned_plot_data)) +
  labs(title = "Compared to what you think of as a typical researcher who uses preexisting observational\ndata in your field, how concerned are you with\nresearch trustworthiness, bias, rigour, and reproducibility ...") +
  theme(
    title = element_text(size = 14),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.title.x = element_text(size = 15),
    legend.text = element_text(size = 10),
    legend.title = element_blank()
  )
```
